---
title: "Using R Markdown"
author: "Dom"
date: 2018-06-16
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
This is some clever stuff. I was first introduced to it by a co-worker who mentioned **Jupyter Notebooks** (another nifty tool) and I didn't see the huge appeal and the applications of it until recently. The same co-worker wanted a R script documented and put into words how it was done.  

I thought why not use the comments amongst the code and go from there but not everyone will want to read through code and possibly have to scroll back through the console looking for outputs amongst the processes performed on the data. That's where **R Markdown** comes in. Time to make this code look somewhat professional and more importantly tell the **story/journey** of the data.  

### Data stories/journies
I keep hearing about telling a story/journey of the data but couldn't figure out when someone would want to hear about it/look through it. Understanding the data story/journey doesn't just apply to executives and higher ups but also other data wranglers but on different levels.

* **Executives/higher ups = less technical story/journey, higher level**
* **Data wranglers = medium-higher techincal story/journey, higher and lower level**

I think that some people can get carried away when discussing the results to higher ups and get tempted into going into too much detail. Yes some smart data prepartion, modelling and predicting may be occuring to get these results but the higher ups want to be able to **make decisions** from the data. 



```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
